---
title: "RV217 - Exploratory Stats"
author: "Bryan Mayer"
date: "`r Sys.Date()`"
output: 
  html_document:
    keep_md: true
    toc: true
    code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


<!---
Creation Date: 07/10/2019
Creation Author: Bryan Mayer
--->



# Setup

```{r load-data}
library(tidyverse)
theme_set(theme_bw())

# num2  keeps APTIMA missing when it didnt have a value 

rv217 = read_csv("../../data/RV217Clean.csv") %>%
  mutate(APTIMA_num2 = case_when(
    APTIMA_num == as.numeric(APTIMA) ~ APTIMA_num,
    TRUE ~ NA_real_
  ))

```


# Repeating results from Robb et al. 2016

- Comparing calculated data to supplemental table
- Reproducing their results
  - Estimating upslope, downslope, set point
  - Repeating correlations

# When was APTIMA measured



# APTIMA vs. VL

Moderate correlation using any viral load but very highly correlated using first positive or dx only. When taking the ratio of the measures, they are almost perfectly correlated specifically at first positive viral load.

```{r vl-vs-aptima}

cor_apt_vl = rv217 %>%
  select(ID, VL, log10VL, APTIMA_num2) %>%
  na.omit() %>% # complete case subsetting but not actually necessary
  mutate(ID = factor(ID)) %>%
  gather(vl_trans, vl, VL, log10VL) %>%
  group_by(vl_trans) %>%
  nest() %>%
  mutate(
    rmcorr_mod = map(data, ~rmcorr::rmcorr(ID, vl, APTIMA_num2, .x)),
    rmcorr_res = map(rmcorr_mod, ~tibble(r = .x$r,pvalue = .x$p, 
                                          CI_lower = .x$CI[1], CI_upper = .x$CI[2])
    )) %>%
  unnest(rmcorr_res)
  

ggplot(data = rv217, aes(x = VL, y = APTIMA_num)) +
  scale_x_log10("VL") +
  scale_y_continuous("APTIMA") +
  geom_point() +
  annotate("text", x = Inf, y = Inf, label = paste0("pho = ", round(cor_apt_vl$r[2], 2)),
           vjust = 1, hjust = 1)

```


```{r dx-vs-firstpos}

rv217_firstpos = rv217  %>%
  group_by(ID, site) %>%
  summarize(
    APTIMA_dx = APTIMA_num2[days_dx == 0],
    APTIMA_firstVL = APTIMA_num2[days == 0],
    firstVL = VL[days == 0]
  ) %>%
  gather(APTIMA_time, APTIMA_value, APTIMA_dx, APTIMA_firstVL)

first_pos_cor = rv217_firstpos %>%
  group_by(APTIMA_time) %>%
  nest() %>%
  mutate(
    spearman_cor = map_dbl(data, ~cor(log10(.x$firstVL), .x$APTIMA_value, 
                                       method = "spearman", use = "complete.obs")),
    spearman_p = map_dbl(data, ~coin::pvalue(coin::spearman_test(
      APTIMA_value ~ log10(firstVL), data = .x)))
    )


rv217_firstpos %>%
  ggplot(aes(x = firstVL, y = APTIMA_value)) +
  scale_x_log10("First pos. VL") +
  scale_y_log10() +
  geom_point() +
  geom_text(data = first_pos_cor, 
            aes(label = paste("pho =", round(spearman_cor, 2)),
                x = Inf, y = Inf, vjust = 1, hjust = 1)) +
  facet_wrap(~APTIMA_time, strip.position = "left", scales = "free_y",
             labeller = as_labeller(c("APTIMA_dx" = "APTIMA at diagnosis",
                                      "APTIMA_firstVL" = "APTIMA at first pos. VL"))) +
  theme(
    axis.title.y = element_blank(),
    strip.placement = "outside",
    strip.background = element_blank(),
    strip.text = element_text(size = 12)
  ) +
  geom_smooth()

```


```{r comp-mod}
mod = rv217_firstpos %>%
  spread(APTIMA_time, APTIMA_value) %>%
  subset(APTIMA_firstVL > 2) %>%
  lm(log(firstVL) ~ log(APTIMA_firstVL), data = .)

```



```{r}

rv217_firstpos %>%
  spread(APTIMA_time, APTIMA_value) %>%
  ggplot(aes(x = firstVL, y = APTIMA_firstVL, colour = site)) +
  scale_x_log10("First pos. VL") +
  scale_y_log10("APTIMA value at first pos. VL") +
  geom_point() +
  theme(
    strip.placement = "outside",
    strip.background = element_blank(),
    strip.text = element_text(size = 12)
  )

```
