---
title: "RV217 - Data Cleaning"
author: "Bryan Mayer"
date: "7/22/2019"
output: 
  html_document:
    keep_md: true
    toc: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


The following script cleans and tidies the raw RV217 data.

**Some notes:**

  - `days` was recalculated to be from the time of the first positive viral load. 
    - Original days variable is now called `days_dx` for days from diagnosis (APTIMA  detection).
  - Used Dan's variable name conventions so our datasets have similar headers
  - No long removing missing IDs, updated processing to handle this better
  - Added a variable for arv. arv day is relative to first positive viral load.

**Some issues:**

- Not every participant has site information
- Not sure what many of the variables mean
- APTIMA needs an LLoQ? Currently: 
    - NR, R or R/NR = NA
- Date variable was used as needed but not generally cleaned for IDs without issues and without ARV (`draw_date`)

# Load data and clean names

```{r load-data}
library(tidyverse)
library(kableExtra)

fix_names = function(x){
  #this loads the header exactly
  suppressMessages({
    x = read_csv("../../data/RV217Master.csv", n_max = 0) %>%
      names()
    })    
  
  x %>%
    str_replace_all("#", "") %>%
    str_replace_all("%", "pct") %>%
    tidy_names(syntactic = T, quiet = T) 
}


# read.csv loads the data than read_csv
rv217_raw = read.csv("../../data/RV217Master.csv", stringsAsFactors = F,
                 na.strings = "") %>%
  as_tibble(.name_repair = fix_names) %>%
  #using Dan's naming conventions
  rename(
    ID = PIN,
    draw_date = Draw.Date,
    VL = Abbott.RealTime.HIV.1.RNA..Copies.ml.,
    log10VL = Abbott.RealTime.HIV.1.RNA.Log..Copies.ml.,
    VL_site = VL.run.by.MHRP..M..or.Site..S.,
    visit_code = Visit
  ) %>%
  mutate(
    VLunit = "copies/mL"
  )

glimpse(rv217_raw)

```

# Error checking

## Dealing with missing IDS

Missing PIN seems to come from populated notes columns

- no cases of viral load or cell data

```{r missing-ids}

missing_id = rv217_raw %>%
  subset(is.na(ID))

nrow(missing_id)

select(missing_id, contains("VL"), contains("CD"), contains("B.")) %>%
  na.omit %>%
  nrow()

```

- There are **three** rows without any data so removing them.

```{r all-missing}
full_missing_rows = map(1:nrow(rv217_raw), function(i){
  if(all(is.na(select(rv217_raw, -VLunit)[i, ]))) i
}) %>%
  flatten_dbl()

full_missing_rows

x = nrow(rv217_raw)
x
rv217_raw = rv217_raw[-c(full_missing_rows), ]
nrow(rv217_raw)
if(x - nrow(rv217_raw) != 3) stop("total missing rows has changed") else rm(x)

# confirm
map(1:nrow(rv217_raw), function(i){
  if(all(is.na(select(rv217_raw, -VLunit)[i, ]))) i
}) %>%
  flatten_dbl()

```

- Identifying which missing IDs are within a series of the same ID

If they are, we can just use locf (assuming master is sorted), otherwise it has to be dealt with manually. There were three cases and in each of these cases, the relevant note/visit code aligns with using locf.

```{r}
tmp_missing_id = map_df(which(is.na(rv217_raw$ID)), function(i){
  rows = c(i - 1, i, i + 1)
  tibble(
    missing_row = i,
    check_rows = str_c(rows, collapse = ","),
    check_id = paste(map_dbl(rows, ~(rv217_raw$ID[.x])), collapse = ","),
    same = str_split_fixed(check_id, ",", n = 3)[,1] == str_split_fixed(check_id, ",", n = 3)[,3] 
  )
})

tmp_missing_id

tmp_missing_id %>%
  subset(!same) %>%
  select(check_rows) %>%
  unlist() %>%
  paste(collapse = ",") %>%
  str_split(",") %>%
  unlist() %>%
  as.numeric %>%
  slice(rownames_to_column(rv217_raw), ., .preserve = T) %>%
  select(rowname, ID, draw_date, Notes, visit_code, days.from.1st.pos)

```

```{r}

rv217_raw = rv217_raw %>%
  mutate(ID = zoo::na.locf(ID))
  

rv217_raw %>%
  subset(is.na(ID)) %>%
  nrow()

```

## Variable types

Checking types of our important endpoints (viral load and cell counts). Random character string in log10VL

```{r endpoint-types}

rv217_raw %>%
  select(contains("VL"), contains("CD"), contains("B.")) %>%
  summarize_all(class) %>%
  gather()

map(rv217_raw$log10VL, function(i){
  if(!is.na(i) & is.na(as.numeric(i))) i
}) %>%
  flatten_chr()

rv217_raw$log10VL[which(rv217_raw$log10VL == "Aptima Reactive")] = NA
rv217_raw$log10VL = as.numeric(rv217_raw$log10VL)

rv217_raw %>%
  select(contains("VL"), contains("CD"), contains("B.")) %>%
  summarize_all(class) %>%
  gather()

```

# Variable munging

##  Priority, Site

There is one unique value per participant. Not sure what prioirty is but site is important so will clean that into Uganda or Thailand unless it is missing.

```{r}
priority_site_dat = rv217_raw %>% group_by(ID) %>%
  summarize(
    total_priority = n_distinct(na.omit(Priority)),
    priority = paste0(na.omit(Priority), collapse = ""),
    total_site = n_distinct(na.omit(Site)),
    site_raw = paste0(na.omit(Site), collapse = "")
  )

priority_site_dat %>%
  subset(total_priority > 1 | total_site > 1) %>%
  nrow()

unique(priority_site_dat$priority)
unique(priority_site_dat$site_raw)

priority_site_dat = priority_site_dat %>%
  mutate(site = case_when(
    str_detect(tolower(site_raw), "th") ~ "Thailand",
    str_detect(tolower(site_raw), "ug") ~ "Uganda",
    TRUE ~ NA_character_
  ))

unique(priority_site_dat$site)
with(priority_site_dat, table(site_raw, site, useNA = "ifany"))

priority_site_dat %>%
  subset(is.na(site)) %>%
  select(ID, site) 

```


```{r merge-site-priority}

rv217_raw = rv217_raw %>%
  select(-Priority, -Site) %>%
  left_join(select(priority_site_dat, ID, priority, site), by = "ID") %>%
  select(-visit_code, -Notes, -EIA, -WB, everything()) # moves misc. variables to the end
  
```

## APTIMA cleaning

Make a variable that is measurement only truncated at minimum (for R and R/NR), truncated at 0 for NR and left missing for weird inputs

```{r aptima}
aptima_vals = map_chr(rv217_raw$APTIMA, function(i){
  if(!is.na(i) & is.na(suppressWarnings(as.numeric(i)))) return(i) else return(NA_character_)
}) 

aptima_vals %>% unique()

rv217_raw = rv217_raw %>%
  mutate(
    APTIMA_num = suppressWarnings(as.numeric(APTIMA))
    )
  
```

# days variable

## Days variable error checking

There are several cases with missing day variables. Can these be calculated?

```{r days-errors}
missing_time_all = rv217_raw %>%
  group_by(ID) %>%
  mutate(
    any_missing_day = any(is.na(days.from.1st.pos))
  ) %>%
  subset(any_missing_day)

# missing times but has date
date_missing_time = missing_time_all %>%
  group_by(ID) %>%
  mutate(
   vl_missing_time = any(is.na(days.from.1st.pos) & !is.na(draw_date))
  ) %>%
  subset(vl_missing_time)

ids_missing_days = unique(date_missing_time$ID)

date_missing_time %>%
  group_by(ID) %>%
  summarize(
    any(days.from.1st.pos == 0)
    )

```

```{r handling-specific-time}

# recode a weird entry
date_missing_time$draw_date[date_missing_time$draw_date == "NO SAMPLES"] <- NA

clean_vl_missing_time = date_missing_time %>%
  group_by(ID) %>%
  mutate(
    infection_time = if(any(!is.na(days.from.1st.pos))) na.omit(draw_date[days.from.1st.pos == 0]) else{
                             na.omit(draw_date[APTIMA == "R"])},
    days.from.1st.pos = as.numeric(difftime(as.Date(draw_date, format = "%d-%b-%y"),
                                 as.Date(infection_time, format = "%d-%b-%y"), 
                                 units = "days"))
  ) %>%
  select(-infection_time, -any_missing_day, -vl_missing_time)


```

## Days variable creation

```{r}
nrow(rv217_raw)

rv217_raw = rv217_raw %>%
  subset(!ID %in% ids_missing_days) %>% # these are processed separately
  bind_rows(clean_vl_missing_time) %>% # and added back here
  rename(days_dx = days.from.1st.pos) %>%
  group_by(ID) %>%
  mutate(
    posVL = !is.na(VL),
    days = days_dx - min(days_dx[posVL])
  ) 

nrow(rv217_raw)

```

# ARV and Primary Kinetics


## ARV

Defining primary kinetics as first pos VL -> min(first ARV, 365 days).

All ARV information is in the Notes. There were three cases where the ARV note text date did not match the ARV note date. I found this through manual inspection and denoted them in the outputted sheet.


```{r}
ARV_notes = rv217_raw %>% 
  subset(str_detect(Notes, "ARV")) %>%
  select(ID, draw_date, days, Notes) %>%
  add_count() %>% 
  group_by(ID) %>%
  mutate(first_note_day = min(days)) %>%
  subset(days == first_note_day) %>%
  rename(arv_note = Notes) %>%
  select(ID, draw_date, days, arv_note) %>%
  mutate(
    arv_matches_date = !ID %in% c(10742, 10204, 40700),
    arv_date = case_when(
      arv_matches_date ~ draw_date,
      ID == 10742 ~ "12-Mar-14",
      ID == 10204 ~ "03-Aug-12",
      ID == 40700 ~ "29-Apr-14",
      TRUE ~ NA_character_
      ),
    adj_time = as.numeric(difftime(as.Date(draw_date, format = "%d-%b-%y"),
                                 as.Date(arv_date, format = "%d-%b-%y"), 
                                 units = "days")),
    arv_day = days - adj_time
    )

ARV_notes

write_csv(ARV_notes, "ARV_notes.csv")

```

## Primary kinetics

```{r}
rv217_raw = ARV_notes %>%
  select(ID, arv_day) %>%
  right_join(rv217_raw, by = "ID") %>%
  mutate(
    arv_note = str_detect(Notes, "ARV"),
    any_arv = any(is.na(arv_day)),
    primary_kinetics = days >= 0 & days <= 365 & (is.na(arv_day) | days < arv_day)
  )

```

# Finalize

Order and save

```{r save-data}
names(rv217_raw) = str_remove_all(names(rv217_raw), "\\.")
glimpse(rv217_raw)


rv217_raw %>%
  select(ID, draw_date, days, VL, log10VL, VLunit, posVL, primary_kinetics, CD4, CD8, NK, B,
         APTIMA, APTIMA_num, days_dx, contains("pct"), everything()) %>%
  write_csv("../../data/RV217Clean.csv")


```